---
title: "power lifeters data_set"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
library(tidyverse)
```

```{r}
ipf_lifts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-08/ipf_lifts.csv")
```

### Inspect NAs in the data : 

```{r}
ipf_lifts %>%
  sapply(is.na) %>%
  colSums()
```

### Curiosity questions: DR style.

```{r}
# what do we exactly mean with division.?
ipf_lifts %>%
  count(division, sort = T)
# A == SI
```

```{r}
# do we have repetetive names.?
ipf_lifts %>% 
  count(name, sort = T) 
# A == SI
```

### data prerp:

I do not think that the place data is relevant beyond 10, so I will be limiting , the analysis to place >= 10

```{r}
# data to analyse: the top 10 performers in each cycle.
data_1 <- ipf_lifts %>% 
  select(- federation) %>%
  filter(place %in% c(1:10)) 
```

The following chunk contains some data cleaning operations which we built up block by block according to the need in each step of the analysis.

```{r}
# remove NAs from division
data_2 <- data_1 %>% 
  filter(!is.na(division)) %>% 
  select(-meet_name)

# rencode the excercice colnames
colnames(data_2)[10:12] <- c("squat", "bench", "deadlift")
# another non elegant way of doing this : 
    #data_3$best <- data_3 %>% 
    #  pull(best) %>% 
    #  substr(6,10)

# encode all character columns as factors, except column name (not elegant)
data_2 <- data_2 %>%
  mutate_if(is.character, as.factor) %>%
  mutate(name = as.character(name)) 

#levels(data_2 %>% pull(place)) <- fct_relevel(data_2$place, "10", after = Inf) 
# something went wrong here: the 4 last levels where integrated with the 6 first, in a way that the sum did not cange
##||##
##VV##
# solution DR style 
places <- c("1","2","3","4","5","6","7","8","9","10")
data_2 <- data_2 %>% mutate(place = fct_relevel(place, places)) # elegent solution
```

to ease analysis I am going to gather the 3 competion excercices into 1 column

```{r}
data_3 <- data_2 %>%
  gather(bench, squat, deadlift, key= "best", value = "kg")
```

Now we can easily visualize and analyze each excercice

```{r}
data_3 %>%
  count(event)
data_3 %>%
  count(equipment)
data_3 %>%
  count(place)
```


```{r}
data_3 %>%
  ggplot(aes(x = kg, fill = event)) + 
  geom_histogram() + 
  facet_grid(sex ~ best) +
  labs(title = "does sex, and event, affect the weight lifted.?")

data_3 %>%
  ggplot(aes(x = equipment, y = kg, fill = sex)) + 
  geom_boxplot() + 
  facet_grid(~ best) + 
  coord_flip() +
  labs(title = "how does the use of equipement differs in terms of sex and weight lifted.?")

data_3 %>%
  ggplot(aes(x = equipment, y = kg, fill = event)) + 
  geom_boxplot() + 
  facet_grid(~ best) + 
  coord_flip() +
  labs(title = "how does the equipement use differs in terms of event and weight lifted.?")
  

data_3 %>%
  ggplot(aes(x = event, y = kg, fill = sex)) + 
  geom_boxplot() +  
  facet_grid(~ best) + 
  labs(title = "how does the weight lifted by sex differs by event.?")

data_2 %>%
  ggplot(aes(x = division, y = deadlift)) + 
  geom_boxplot() + 
  facet_grid(sex~.) +
  coord_flip() +
  labs(title = "how does the division lifters differ by sex")

data_2 %>%
  select(division, deadlift, bench, squat) %>%
  group_by(division) %>%
  summarise(deadlift = mean(deadlift),
            squat = mean(squat),
            bench = mean(bench)) %>%
  gather(deadlift, squat, bench, key = "best", value = "mean") %>%
  ggplot(aes(x = division, y = mean)) + 
  geom_col() + 
  facet_grid(~best) + 
  coord_flip()

data_2 %>%
  select(division, deadlift, bench, squat) %>%
  group_by(division) %>%
  summarise(deadlift = median(deadlift),
            squat = median(squat),
            bench = median(bench)) %>%
  gather(deadlift, squat, bench, key = "best", value = "median") %>%
  ggplot(aes(x = division, y = median)) + 
  geom_col() + 
  facet_grid(~best) + 
  coord_flip()
```


```{r}
data_3 %>%
  select(division, place, best, kg) %>%
  ggplot(aes(x = division, y = kg, colour = best)) +
  geom_point(alpha = 0.5) + 
  coord_flip() +
  facet_wrap(vars(place)) +
  labs(title = "how are the places distributed by division to the weight lifted and to the excercice")

data_3 %>%
  select(division, place, sex) %>%
  count(place, division, sex) %>%
  ggplot(aes(x = place, y = n, fill = sex)) +
    geom_col()+
    facet_wrap(vars(division), scales = "free") 

```


**To deepen the analysis, I think hypthesis testing is the go to solution, to test the significance of being in a group and actually being in the top 3**





